\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2017-01-05 Response.tex

\Cref{prop:eR_dot_bound} has been modified to include additional detail on the upper bound of \( A(R) \) as:
\begin{correction}
    Upper bounds of \( A(R) \) and \( B(R) \) are given by
    \begin{gather}
        \norm{A} < b_2 \norm{e_{R_A}}^2 = c_A  , \quad \norm{B} < c_B . 
    \end{gather}
  where the constant \( b_2\) is given by \( b_2 = \frac{h_1 h_4}{h_5 \parenth{h_1 - \psi}}\) for
  \begin{align*}
    h_4 &= \min\braces{g_1 + g_2, g_2 + g_3 , g_3 + g_1} ,\\
    h_5 &= \min\braces{\parenth{g_1 + g_2}^2,\parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2} .\\
  \end{align*}
\end{correction}

The domain \( D \) is chosen to both exclude the undesired critical points of \( A(R)\) as well as the infeasible regions defined by the constraints. 
In general, this is not a difficult assumption to satisfy as the desired attitude will typically be located far from the constrained regions.
We have augmented the manuscript with some additional discussion of the chosen domain \( D \) in~\Cref{proof:eR_dot_bound}.

\begin{correction}
    We choose the domain \( D \) to exclude the undesired critical points of \( A(R) \) as well as the infeasible regions defined by the state inequality constraints. 
    Furthermore, one can show that the set \( D \) is an open neighborhood of the desired attitude configuration, \( R_d\).
    The proof of the upper bound of \( A(R) \) is given in~\cite{LeeITCST13}.
    The selected domain ensures that the configuration error function is bounded \( \Psi < \psi \).
    This implies that that both \( A(R) \) and \( B(R) \) are bounded by constants \( c_A c_B < \psi < h_1\).
    Furthermore, since \( \norm{B} > 1 \) this ensures that \( c_A, c_B < \psi\) and shows~\cref{eqn:AB_bound}.
\end{correction}

The proof of~\Cref{prop:adaptive_control} given in~\Cref{proof:adaptive_control} has been improved. 
Originally, the positive definiteness of \( \mathcal{V}\) was not explicitly defined. 
In order to prove that the Lyapunov function is positive definite an additional quadratic property of the configuration error function was derived.
A function \( \psi \) is quadratic on the domain \( D \) about the critical zero \( x_0 \in D \) if there exists two constants \( 0 < b_1 \leq b_2 \) such that
\begin{align*}
    b_1 \norm{\text{d} \psi}^2 \leq \psi(x) b_2 \leq \norm{\text{d} \psi}^2 ,
\end{align*}
for all \( x \in D \, \setminus \, x_0\)~\cite[Definition 6.28]{bullo2004}.
To show that \( \Psi \) is quadratic we compute the hessian of \( \Psi \) at the desired attitude \( R = R_d \).
This additional detail was included in~\cref{prop:config_error}, as shown here:
\begin{correction}
     \( \Psi \) is a locally quadratic function on the domain 
        \begin{align}\label{eq:psi_quadratic_domain}
            D_1 = \braces{R \in \SO \vert \Psi = 0, e_{R} = 0 \iff R = R_d } ,
        \end{align}
        and it is bounded by constants \( 0 < n_1 \leq n_2 \) as
        \begin{align}\label{eq:psi_bound}
            n_1 \norm{e_R}^2 \leq \Psi(R) \leq n_2 \norm{e_R}^2 .
        \end{align}
and proven in~\Cref{proof:config_error}, as shown here:
\begin{correction}
    We show~\cref{item:prop_psi_quadratic} by computing the hessian of \( \Psi \), \( \Hess{\Psi} \),  using the chain rule as 
    \begin{align*}
        \Hess \Psi =& \parenth{\mathbf{D}^2 A \cdot \delta^2 R} B + \parenth{\dirDiff{A}{R} }\parenth{\dirDiff{B}{R}}^T \\
        &+ \parenth{\dirDiff{B}{R} }\parenth{\dirDiff{A}{R}}^T +\parenth{\mathbf{D}^2 B \cdot \delta^2 R} A .
    \end{align*}
    The first order derivative of \( A(R) \) and \( B(R) \) are given by~\cref{eqn:eRA,eqn:eRB}. 
    The hessian of \( A(R) \) is computed as 
    \begin{align*}
        \dirDiff{\parenth{\dirDiff{A}{R}}}{R} &= \eta \cdot \frac{1}{2}\parenth{G R_d^T R \delta R - \delta R^T R_d G}^\vee , \\
        &= \eta \cdot \frac{1}{2} \parenth{G R_d^T R \hat \eta + \hat \eta R^T R_d G}^\vee,\\
        &= \eta \cdot \frac{1}{2} \bracket{\parenth{\braces{\tr{R^T R_d G} I - R^T R_d G} \eta}^\wedge}^\vee , \\
        &= \eta \cdot \frac{1}{2} \parenth{\tr{R^T R_d G} I - R^T R_d G}\eta ,
    \end{align*}
    where we used the scalar triple product rule, from~\cref{eqn:STP}, to arrive at the final form.
    The hessian of \( B(R) \) is computed as
    \begin{align*}
        \dirDiff{\parenth{\dirDiff{B}{R}}}{R} =& \eta \cdot \left[\frac{\parenth{\delta R^T v}^\wedge r}{\alpha\parenth{r^T R^T v - \cos \theta}} \right. \\
        & \left. - \frac{\alpha \parenth{R^T v}^\wedge r \parenth{r^T \delta R^T v}}{\alpha^2 \parenth{r^TR^Tv - \cos \theta}^2}   \right] .
    \end{align*}
    The term \( \parenth{\delta R^T v}^\wedge r\) is simplified as
    \begin{align*}
        \parenth{\delta R^T v}^\wedge r &= - \hat{r} \delta R ^T v , \\
        &= - \hat{r} \parenth{R \hat \eta}^T v , \\
        &= \hat{r} \parenth{\hat{\eta} R^T } v , \\
        &= - \hat{r} \parenth{R^T v}^\wedge \eta ,
    \end{align*}
    where we utilized the hat map property from~\cref{eqn:cross_product}.
    Similarly, the term \( r^t \delta R^T v \) is simplified as
    \begin{align*}
        r^T \delta R^T v &= r^T \parenth{-\hat \eta R^T}v , \\
        &= r^T \parenth{R^T v}^\wedge \eta .
    \end{align*}
    The hessian of \( B(R) \) then becomes
    \begin{align*}
        \dirDiff{\parenth{\dirDiff{B}{R}}}{R} =& \eta \cdot \left[ - \frac{ \hat r \parenth{R^T v}^\wedge}{\alpha \parenth{r^T R^T v - \cos \theta}} \right. \\ 
        & \left. - \frac{\parenth{R^T v}^\wedge r r^T \parenth{R^T v}^\wedge }{ \alpha\parenth{r^T R^T v - \cos \theta}^2}  \right] \eta .
    \end{align*}
    Using these terms, we evaluate \( \Hess{\Psi} \) at the desired attitude \( R = R_d \) as
    \begin{align*}
        \Hess{\Psi} = \eta^T \frac{1}{2} B \parenth{\tr{G}I -  G} \eta , 
    \end{align*}
    which is positive definite since \( B > 1\) and \( \sum g_i > g_i\). 
    From~\cite[Proposition 6.30]{bullo2004}, \( \Psi \) is locally quadratic on the set \( D_1 \) which proves~\cref{item:prop_psi_quadratic}.

Subsequently,~\Cref{prop:adaptive_control} was modified to explicitly demonstrate that \( \mathcal{V} \) is positive definite.
\begin{correction}
    Consider the Lyapunov function \( \mathcal{V} \) given by
    \begin{align*}
    \mathcal{V} = \frac{1}{2} e_\Omega \cdot J e_\Omega + k_R \Psi + c J e_\Omega \cdot e_R + \frac{1}{2 k_\Delta} e_\Delta \cdot e_\Delta , %\label{eqn:v_adapt}
    \end{align*}
    over the domain \( D_3 = D_1 \cap D_2\), which ensures that the properties of~\Cref{prop:config_error,prop:error_dyn} are satisfied.

    From~\Cref{prop:config_error}, the configuration error function is locally quadratic and it is bounded in \( D_3 \) by~\cref{eq:psi_bound}~\cite[Definition 6.28]{bullo2004}.
    Using this the Lyapunov function \( \mathcal{V} \) is bounded by
    \begin{align*} %\label{eqn:v_upper_bound}
    z^T W_1 z \leq \mathcal{V} \leq z^T W_2 z ,
    \end{align*}
    where \( e_\Delta = \Delta - \bar{\Delta} \), \( z = [\|e_R\|,\|e_\Omega\|,\|e_\Delta\|]^T\in\R^3 \) and the matrices \(W_1,W_2 \in \R^{3 \times 3}\) are given by
    \begin{align*}
    W_1 & = \begin{bmatrix}
        k_R n_1 & \frac{1}{2} c \lambda_m & 0 \\
        \frac{1}{2} c \lambda_m & \frac{1}{2} \lambda_m & 0 \\
        0 & 0 & \frac{1}{2 k_\Delta}
    \end{bmatrix},\\
    W_2 & = \begin{bmatrix}
        k_R n_2 & \frac{1}{2} c \lambda_M & 0 \\
        \frac{1}{2} c \lambda_M & \frac{1}{2} \lambda_M & 0 \\
        0 & 0 & \frac{1}{2 k_\Delta}
    \end{bmatrix} .
    \end{align*}
    The time derivative of \( \mathcal{V}\) with the control input defined by~\cref{eqn:adaptive_control} is given as
    \begin{align}
    \dot{\mathcal{V}} =& - k_\Omega e_\Omega^T e_\Omega + \parenth{e_\Omega + c e_R}^T W e_\Delta - k_R c e_R^T e_R \nonumber\\
    &- k_\Omega c e_R^T e_\Omega + c J e_\Omega^T \dot{e}_R - \frac{1}{k_\Delta} e_\Delta^T \dot{\bar{\Delta}} , \label{eqn:vdot}
    \end{align}
    where we used \( \dot{e}_\Delta = - \dot{\bar{\Delta}} \).
    The terms linearly dependent on \( e_\Delta\) are combined with~\cref{eqn:delta_dot} to yield
    \begin{align*}
     e_\Delta^T \parenth{W^T \parenth{e_\Omega + c e_R} - \frac{1}{k_\Delta} \dot{\bar{\Delta}}} = 0 . 
    \end{align*}
    Using~\Cref{prop:eR_dot_bound} an upper bound on \( \dot{\mathcal{V}} \) is written as
    \begin{gather*}
    \dot{\mathcal{V}} \leq -\zeta^T M \zeta ,
    \end{gather*}
    where $\zeta=[\|e_R\|,\|e_\Omega\|]\in\R^2$, and the matrix \( M \in \R^{2 \times 2} \) is 
    \begin{gather*}
    M = \begin{bmatrix}
        k_R c & \frac{k_\Omega c}{2} \\
        \frac{k_\Omega c}{2} & k_\Omega - c \lambda_M H
    \end{bmatrix} .
    \end{gather*}
    If \( c \) is chosen such that~\cref{eqn:c_bound} is satisfied then the matrices \( W_1, W_2 \) and \( M \) are positive definite.
    This implies that $\mathcal{V}$ is positive definite and decrescent, and $\dot{\mathcal{V}}$ is negative semidefinite in the domain $D_3$. As such, the zero equilibrium is stable in the sense of Lyapunov, and all of the error variables are bounded. Furthermore, $\lim_{t\to\infty} \zeta=0$ according to the LaShalle-Yoshizawa theorem~\cite{khalil1996}. 
\end{correction}

### 2017-01-05 manuscript.tex
Improve surface plots

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \) toward the desired attitude, and a repulsive term, \( B(R) \) away from the undesired direction \( R^T v \).
In order to visualize the attitude error function on \( \SO \), we utilize a spherical coordinate representation.
Recall, that the spherical coordinate system represents the position of a point relative to an origin in terms of a radial distance, azimuth, and elevation.
This coordinate system is commonly used to define locations on the Earth in terms of a latitude and longitude.
Similarly, the positions of celestial objects are defined on the celestial sphere in terms of right ascension and declination. 
\end{multicols}
\begin{figure}[H]%
    \centering 
    \subfigure[Attractive \( A(R) \) \label{fig:attract_error}]{\includegraphics[width=0.3\columnwidth]{attract_error} }~
    \subfigure[Repulsive \( B(R) \) \label{fig:avoid_error} ]{\includegraphics[width=0.3\columnwidth]{avoid_error} }~
    \subfigure[Combined \( \Psi \) \label{fig:combined_error}]{\includegraphics[width=0.3\columnwidth]{combined_error} }%
    \caption{Visualization of Configuration Error Functions using spherical coordinate representation}
    \label{fig:config_error} 
\end{figure}%
\begin{multicols}{2}
We apply this concept and parametrize the rotation matrix \( R \in \SO \) in terms of the spherical angles \( \SI{-180}{\degree} \leq \lambda \leq \SI{180}{\degree}  \) and \( \SI{-90}{\degree} \leq \beta \leq \SI{90}{\degree} \). 
Using the elementary Euler rotations, the rotation matrix is now defined as \( R = \exp( \lambda \hat{e}_2) \exp( \beta \hat{e}_3) \).
We iterate over the domains of \( \lambda\) and \(\beta\) in order to rotate the body-fixed vector \( r \) throughout the two-sphere \( \S^2 \).
Applying this method,~\cref{fig:config_error} allows us to visualize the error function on \( \SO \).
The horizontal axes of~\cref{fig:config_error} represent the domain of the spherical angles \( \lambda \) and \( \beta \) in degrees, while the vertical axes represent the unitless magnitude of the error functions defined in~\cref{eqn:psi,eqn:A,eqn:B}.
The attractive error function, given by~\cref{eqn:A}, has been previously used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}, where the desired attitude lies at the minimum of \( A(R) \).

>>>>

\Cref{eqn:psi} is composed of an attractive term, \( A (R) \) toward the desired attitude, and a repulsive term, \( B(R) \) away from the undesired direction \( R^T v \).
In order to visualize the attitude error function on \( \SO \), we utilize a spherical coordinate representation.
Recall, that the spherical coordinate system represents the position of a point relative to an origin in terms of a radial distance, azimuth, and elevation.
This coordinate system is commonly used to define locations on the Earth in terms of a latitude and longitude.
Similarly, the positions of celestial objects are defined on the celestial sphere in terms of right ascension and declination. 
\end{multicols}
\begin{figure}[H]%
    \centering 
    \subfigure[Attractive \( A(R) \) \label{fig:attract_error}]{\includegraphics[width=0.33\columnwidth]{attract_error} }~
    \subfigure[Repulsive \( B(R) \) \label{fig:avoid_error} ]{\includegraphics[width=0.33\columnwidth]{avoid_error} }~
    \subfigure[Combined \( \Psi \) \label{fig:combined_error}]{\includegraphics[width=0.33\columnwidth]{combined_error} }%
    \caption{Visualization of Configuration Error Functions using spherical coordinate representation}
    \label{fig:config_error} 
\end{figure}%
\begin{multicols}{2}
We apply this concept and parametrize the rotation matrix \( R \in \SO \) in terms of the spherical angles \( \SI{-180}{\degree} \leq \lambda \leq \SI{180}{\degree}  \) and \( \SI{-90}{\degree} \leq \beta \leq \SI{90}{\degree} \). 
Using the elementary Euler rotations, the rotation matrix is now defined as \( R = \exp( \lambda \hat{e}_2) \exp( \beta \hat{e}_3) \).
We iterate over the domains of \( \lambda\) and \(\beta\) in order to rotate the body-fixed vector \( r \) throughout the two-sphere \( \S^2 \).
Applying this method,~\cref{fig:config_error} allows us to visualize the error function on \( \SO \).
The horizontal axes of~\cref{fig:config_error} represent the domain of the spherical angles \( \lambda \) and \( \beta \) in degrees, while the vertical axes represent the unitless magnitude of the error functions defined in~\cref{eqn:psi,eqn:A,eqn:B}.
The attractive error function, given by~\cref{eqn:A}, has been previously used for attitude control on \(\SO\).
The potential well of \( A(R)\) is illustrated in~\cref{fig:attract_error}, where the desired attitude lies at the minimum of \( A(R) \).
